---
title: "Statistical Inference"
subtitle: "Repeated Measures Anova"
author: "Your name here"
date: Published on `r format(Sys.time(), "%A %d %B %Y")`
output:
  rmdformats::readthedown:
    highlight: kate
    toc_float: true
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(ggpubr)
library(datarium)
library(rstatix)
```

## Introduction

The repeated measures ANOVA makes the following assumptions about the data:

-   **No significant outliers** in any cell of the design. This can be checked by visualizing the data using box plot methods and by using the function `identify_outliers()` [rstatix package].

-   **Normality**: the outcome (or dependent) variable should be approximately normally distributed in each cell of the design. This can be checked using the **Shapiro-Wilk normality test** (`shapiro_test()` [rstatix]) or by visual inspection using **QQ plot** (`ggqqplot()` [ggpubr package]).

-   **Assumption of sphericity**: the variance of the differences between groups should be equal. This can be checked using the **Mauchly's test of sphericity**, which is automatically reported when using the R function `anova_test()` [rstatix package].

Note that, [**if the above assumptions are not met**]{.ul} there are a non-parametric alternative (*Friedman test*) to the one-way repeated measures ANOVA!

Unfortunately, there are no non-parametric alternatives to the two-way and the three-way repeated measures ANOVA. Thus, in the situation where the assumptions are not met, you could consider running the two-way/three-way repeated measures ANOVA on the transformed and non-transformed data to see if there are any meaningful differences.

If both tests lead you to the same conclusions, you might not choose to transform the outcome variable and carry on with the two-way/three-way repeated measures ANOVA on the original data.

It's also possible to perform robust ANOVA test using the **WRS2** R package.

No matter your choice, you should report what you did in your results.

## RM Anova in R

Key R functions:

-   `anova_test()` [rstatix package], a wrapper around `car::Anova()` for making easy the computation of repeated measures ANOVA. Key arguments for performing repeated measures ANOVA:

    -   `data`: data frame

    -   `dv`: (numeric) the dependent (or outcome) variable name.

    -   `wid`: variable name specifying the case/sample identifier.

    -   `within`: within-subjects factor or grouping variable

-   `get_anova_table()` [rstatix package]. Extracts the ANOVA table from the output of `anova_test()`. It returns ANOVA table that is automatically corrected for eventual deviation from the sphericity assumption. The default is to apply automatically the Greenhouse-Geisser sphericity correction to only within-subject factors violating the sphericity assumption (i.e., Mauchly's test p-value is significant, p \<= 0.05).

## 1-way RM Anova

The dataset "***selfesteem***" contains 10 individuals' self-esteem score on three time points during a specific diet to determine whether their self-esteem improved.

```{r}
data("selfesteem", package = "datarium")
head(selfesteem, 3)

```

The one-way repeated measures ANOVA can be used to determine whether the means self-esteem scores are significantly different between the three time points. So let's convert this data frame into long format:

```{r}
selfesteem <- selfesteem %>%
  gather(key = "time", value = "score", t1, t2, t3) %>%
  convert_as_factor(id, time)
head(selfesteem, 3)
```

### Descriptive statistics

```{r}
selfesteem %>%
  group_by(time) %>%
  get_summary_stats(score, type = "mean_sd")
```
```{r}
bxp<- ggboxplot(selfesteem, y="score", x="time") 
bxp
```


### Assumptions


```{r}
selfesteem %>%
  group_by(time) %>%
  identify_outliers(score)
```
```{r}
selfesteem %>%
  group_by(time) %>%
  shapiro_test(score)
```
```{r message=FALSE, warning=FALSE}
ggqqplot(selfesteem, "score") +
  facet_grid(~time)
```

sphericity test error

```{r}


```
### Anova
```{r}
results <- anova_test(data = selfesteem,
           dv= score,
           wid= id,
           within = time)

```



### Post-hoc tests

```{r}


pwc <- selfesteem %>%
  emmeans_test(score ~time, p.adjust.method = "bonferroni")
pwc
```
```{r}
pwc <- pwc %>% add_xy_position(x="time")
bxp + 
  stat_pvalue_manual(pwc) +
  labs(
    subtitle=get_test_label(results,detailed=TRUE),
    caption=get_pwc_label(pwc)
  )

```

### Conclusions

Hence, we can conclude that the score was statistically significantly different during different times, with the result with F(2,18) = 55.5 and p = 0.0000000201. From the post-hoc tests, when pairwise comparisons were conducted we can state that all the differences between pairs are statistically significant.



## 2-way RM Anova

For Two-Way Repeated Measures ANOVA, "Two-way" means that there are two factors in the experiment, for example, different treatments and different conditions. "Repeated-measures" means that the same subject received more than one treatment and/or more than one condition. Similar to two-way ANOVA, two-way repeated measures ANOVA can be employed to test for significant differences between the factor level means within a factor and for interactions between factors.

Using a standard ANOVA in this case [is not appropriate]{.ul} because it fails to model the correlation between the repeated measures, and the data violates the ANOVA assumption of independence. Two-Way Repeated Measures ANOVA designs can be two repeated measures factors, or one repeated measures factor and one non-repeated factor. If any repeated factor is present, then the repeated measures ANOVA should be used.

Please apply Two-way RM-ANOVA to analyze if any of interactions are significant (*between time and music, time and image, music and image, or music and time and image*)! The response variable is level of stress experienced by a person watching one of 2 movie genres. Interpret your results. Use the following data set:

```{r}
set.seed(5250)
myData <- data.frame(PID = rep(seq(from = 1,
                               to = 60, by = 1), 20),
                     stress = sample(x = 1:100,
                                     size = 1200,
                                     replace = TRUE),
                     image = sample(c("Happy", "Angry"),
                                    size = 1200,
                                    replace = TRUE),
                     music = sample(c("Disney", "Horror"),
                                    size = 1200,
                                    replace = TRUE)
)
myData <- within(myData, {
  PID   <- factor(PID)
  image <- factor(image)
  music <- factor(music)
})
myData <- myData[order(myData$PID), ]
head(myData)
```

```{r}

head(myData, 60)

```

The one-way repeated measures ANOVA can be used to determine whether the means self-esteem scores are significantly different between the three time points. So let's convert this data frame into long format:

```{r}
selfesteem <- selfesteem %>%
  gather(key = "time", value = "score", t1, t2, t3) %>%
  convert_as_factor(id, time)
head(selfesteem, 3)
```

### Descriptive statistics

```{r}
selfesteem %>%
  group_by(time) %>%
  get_summary_stats(score, type = "mean_sd")
```
```{r}
bxp<- ggboxplot(selfesteem, y="score", x="time") 
bxp
```


### Assumptions


```{r}
selfesteem %>%
  group_by(time) %>%
  identify_outliers(score)
```
```{r}
selfesteem %>%
  group_by(time) %>%
  shapiro_test(score)
```
```{r message=FALSE, warning=FALSE}
ggqqplot(selfesteem, "score") +
  facet_grid(~time)
```

sphericity test error

```{r}


```
### Anova
```{r}
results <- anova_test(data = selfesteem,
           dv= score,
           wid= id,
           within = time)

```



### Post-hoc tests

```{r}


pwc <- selfesteem %>%
  emmeans_test(score ~time, p.adjust.method = "bonferroni")
pwc
```
```{r}
pwc <- pwc %>% add_xy_position(x="time")
bxp + 
  stat_pvalue_manual(pwc) +
  labs(
    subtitle=get_test_label(results,detailed=TRUE),
    caption=get_pwc_label(pwc)
  )

```
